---
title: "TELEMAC (Topic Explorator Lab of European Media Agenda applied to Crisis)"
subtitle: "Version 2.2 - June 2021"
author: "Claude Grasland, Université de Paris"
output: html_document
---


```{r setup, echo = FALSE, message = FALSE, warning = FALSE, error = FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, error = FALSE)
library(sf)
library(plotly)
library(RColorBrewer)
library(data.table)
library(DT)
library(dplyr)
library(cartography)
library(knitr)
library(tidygraph)
library(ggraph)
library(ggplot2)
#library(reldist)
#library(quanteda)

#load hypercube by weeks
hc <- readRDS("data/hc_EUR_foreign_mobil_week.Rdata")
hc<-hc[hc$when>as.Date("2013-12-31"),]
hc<-hc[hc$when<as.Date("2020-01-01"),]
ha<-hc[is.na(hc$when)==FALSE,]

# store week in a list of hypercubes
hyper<-list()
hyper$week<-hc

#aggregate by month and store
sel<-hc
sel$when<-as.Date(cut(sel$when,breaks =  "month"))
hyper$month<-sel[, list(news=sum(news)),by =list(order,who, when, where1, where2, what)]

#aggregate by quarter and store
sel<-hc
sel$when<-as.Date(cut(sel$when,breaks =  "quarter"))
hyper$quarter<-sel[, list(news=sum(news)),by =list(order,who, when, where1, where2, what)]

#aggregate by month and store
sel<-hc
sel$when<-as.Date(cut(sel$when,breaks =  "year"))
hyper$year<-sel[, list(news=sum(news)),by =list(order,who, when, where1, where2, what)]

```



This notebook present the functions used by the interface TELEMAC presented in the paper published in the journal *Frontiers* and available at https://analytics.huma-num.fr/Claude.Grasland/telemac/ 

# DATA COLLECTION

## Selection of newspapers

We have collected titles of news declared as *international*  (an article is an article where as least one foreign country has been detected by the tagging procedure).
In each country of analysis, we selected newspapers with a national audience for which we have the avaible data for a period of seven years (mid 2013 to mid 2020).

```{r, source}

sel<-names(table(hc$who))

meta<-read.csv("data/meta.csv",
               header=T,
               sep = ";",
               encoding = "UTF-8")
meta<-meta[,c(2,4,3)]


kable(meta, col.names = c("Code","Name","URL"), caption = "Selection of daily newspapers")
```
## Selection of textual units

Each item of the RSS feeds has been divided in a maximum of four sentences (textual units):
* The title is considered as the first sentence,
* A part of the description with a maximum size of three sentences. 
Longer descriptions has been rejected from the data set.
The following table present the the number of textual units (sentences) aviable in each newspaper.
The textual content of two newspapers ( _The Guardian_, _El Mundo_) is limited to the only title.

```{r}
tab <- hc 
tab$order<-as.factor(tab$order)
levels(tab$order)<-c("Title", "1st sentence", "2nd sentence","3rd sentence")
tab <- tab[,.(n = round(sum(news),0)), list(order,who)]
tab<-dcast(tab,who~order, fill=0)
mat<-as.matrix(tab[,-1])
row.names(mat)<-tab$who
tabfin <- addmargins(mat)



kable(tabfin,caption = "Textual units by media and order in news" )
```
## Distribution of news by country and by week

For the full period of observation, the total number of textual units published weekly in each country of the sample is between 500 and 2000.

```{r}
tab <- hc[hc$when < as.Date("2020-06-28"),]
tab$week<-cut(tab$when,breaks = "week")
tab$country<-substr(tab$who,4,6)
tab <- tab[,.(n = round(sum(news),0)), list(week,country)]
tab$week<-as.Date(tab$week)

p <- ggplot(tab, aes(x=week, y = n,colour=country)) +
      geom_line() +
      scale_y_continuous("Number of textual units") +
        scale_x_date("Week")

p


```



# BUILDING AN HYPERCUBE


## Multiple dimensions for the observation of news

The hypercube is the result of an aggregation of foreign news according several dimensions:

-   **who** : this dimension is related to the variable which describe the media outlets which published the RSS feeds.
Each source is related to a code`ll_sss_xxxxxx` where `ll`is the language, `sss`is the ISO3 code of the country and `xxxxxx`the name of the media. 
For instance, a RSS feed produced by the French newspaper *Le Figaro* is identified by the code who = `fr_FRA_figaro`. 
Starting from there, it is then possible to proceed to aggregation of the data by group of languages (eg. computation of the indicators for all the French speaking newspapers) or countries (compute the indicateurs for all the media outlets located in France).

-   **when** : this dimension describe the period when an article of the RSS feeds has been published, according to the time zone of Paris.
We note that this variable was transformed from the original data.: the Mediacloud collect the data according to the time zone of Boston.
In the interface, the data are aggregated according to different period of aggregation: weeks, months, quarters or years. 
In each case, the variable `when` is identified by the date of the first day of the period. 
For instance, by choosing to work on monthly aggregated data, the first period of observation for the year 2015 will be: `when = 2015-01-01`. 
These different aggregated data are pre-computed in the data base used by the TELEMAC interface: it is not possible to access to the daily data published by the newspaper and avaible in the Mediacloud data base.


-   **where1** and **where2** : this dual dimension is associated to the cross-list of foreign countries detected by the country dictionary in the news. 
For example the news ("*Conflict between Russia and Turkey about Syria*") will lead to list of three places (*RUS,TUR,SYR*) associated to the cross-list of nine couple of places (*RUS-RUS, RUS-TUR, RUS-SYR, TUR-RUS, TUR-TUR, TUR-SYR, SYR-RUS, SYS-TUR, SYR-SYR*) where each couple will receive a weight of 1/9. 
It is important to keep in mind that the countries where the media are located (mentionned in the `who` dimension) are excluded from the list (see our ambition to work foreign news). 
We excluded fro the data set the news where any countries is mentioned.

-   **what** : this dimension refers to the topic of interest for the exploration. The topic is identified through a boolean (presence of the topic: `what = _yes_`, absence of the topic: `what = _no_` )
This solution is applied 1) for the border topic: the variable identify the articles where the keyword "border" is detected in the language of the source (using a multilingual dictionary for the tagging procedure). The topic of international migration is composed by several group of words related to three different way to speak about immigrant people : "migrants", "refugee" and "asylum seekers" (`what = MIG, what = REF, what = ASY`) which refers to news that use the lexical terms of *refugees* or *migrants* or *ayslum seekers* (again: this procedure use a multilingual dictionary taking in account the variation of vocabulary specific of each national context).
When two subtopics are detected in the same text, the weigh of the news is shared between the different topics. For example a news entitled (*"Migrant and refugees from Syria arrived in Hungary"*) will be broken in 8 cells of the hypercube, each of them with weigh 1/8th corresponding to the combinations (*MIG-SYR-SYR, MIG-SYR-HUN, MIG-HUN-SYR, MIG-HUN-HUN, REF-SYR-SYR, REF-SYR-HUN, REF-HUN-SYR, REF-HUN-HUN*) because we have to combine the topic dimension `what` with the previous dimensions `where1`and `where2`.

To build the hypercube, it is possible to works on different size of text units: (`order=1`): the title or the first sentence, or (`order = 2,3,4, ...`): the title with the selected number of sentence of the description abaible
This parameter is important because some results, especially regarding the spatial dimension of the analysis (where) are more noticeable on longer texts.



## Dictionary of countries

The aim is to create a World map of states which includes an extended list of UN Members, and all pieces of land belonging to a state, whatever its legal status.The states list gathers:

-   the 193 member states of the UN,
-   the 2 observer states of the UN : Palestine and the Holy See (Vatican),
-   the states recognized by at least one member of the UN: Kosovo, Taiwan, Western Sahara, Abkhazia, South Ossetia, Northern Cyprus.

The base map of the 201 spatial units is extracted from [Natural Earth](https://www.naturalearthdata.com/) with a high level of resolution for the location of the centroid of all spatial units (`world_ctr`). But a simplified version is produced with the geometry of the 177 larger units only for quicker visualization in the graphic interface (`world`). Both files are stored in spatial features format from the package R `sf` which is currently the most practical storage. The default projection is the Winkel Tripel, centered on the meridian of Greenwich.

```{r worldmap, echo = FALSE}
world_ctr<-readRDS("map/world_ctr.Rdata")
world_map<-readRDS("map/world_map.Rdata")



par(mar=c(2,0,3,0), mfrow=c(1,1))
typoLayer(world_map,var ="ISO3", 
          legend.pos = "n",
          border = "grey40",
          lwd = 0.5)
plot(world_ctr$geometry, cex=0.3,col = "red", add=TRUE)

layoutLayer(title = "The world in 201 states - sovereign, observer or partly recognized by United Nation", 
            scale = "auto",
            sources = "Natural Earth, proj. Winkel-Tripel")

```


The recognition of states is based on a limited of keywords or regular expression which concern the name of the country, the adjective and/or the gentile of inhabitants and the capital city. The regular expression are used cautiously in order to exclude false positive. 

```{r, dico states}

sel<-names(table(hc$who))

dico<-read.csv("dict/keywords_states_apr_2021.csv",
               header=T,
               sep = ";",
               encoding = "UTF-8")
sel<-dico %>% filter(lang %in% c("en","fr","de","es","it")) %>% filter(ISO3=="SYR")


kable(sel, col.names = c("ISO3","regexp","language"), caption = "Sample of the dictionary of countries (Syria)")
```

## Dictionary of border topic

The dictionary of border topic is voluntary limited to one or two words or regular expression in each language related to the political divisions of the world. We exclude words that are more ambiguous and are not necessary related to geopolitical divisions like *frontier* or *limit* in english.


```{r, dico border}

sel<-names(table(hc$who))

dico<-read.csv("dict/keywords_border_apr_2021.csv",
               header=T,
               sep = ";",
               encoding = "UTF-8")
sel<-dico %>% filter(lang %in% c("en","fr","de","es","it"))


kable(sel, col.names = c("ISO3","regexp","language"), caption = "Dictionary of border topic")
```


## Dictionary of international human mobility

The dictionary of international human mobility is more complex as it is based on two different concept which correspond in english to *migrant* and *refugees* and which can be observed in the other languages under investigation. The first concept (designed as **MIG**) is more general but is also often associated to more negative visions of human mobility across borders. The second concept (designed as **REF**) is generally related to legal issues but can also be associated to "politically correct" way of description associated to a more positive - or less negative- vision of human mobility. These dual concepts insure a bette coverage of the topic in different languages but can also be used to discover semantic shifts inside the same language. For a more detailed discussion, see. [Leconte R., Toureille E. & Grasland C., 2020](https://journals.openedition.org/socio-anthropologie/6235)  


```{r, dico mobil}

sel<-names(table(hc$who))

dico<-read.csv("dict/keywords_mobil_apr_2021.csv",
               header=T,
               sep = ";",
               encoding = "UTF-8")
sel<-dico %>% filter(lang %in% c("en","fr","de","es","it"))


kable(sel, col.names = c("ISO3","regexp","language"), caption = "Dictionary of mobility topic")
```


## Storage and aggregation ofthe hypercube 

Considering the potential size of the hypercubes, we have chosen an efficient format of storage with the R package `data.table` (<https://rdatatable.gitlab.io/data.table/>) which is recognized as more efficient for large computation than the classical `data.frame` or `tibble` formats.

In order to reduce the size of storage we do not store the empty cells of the cube but who have to keep in mind the fact that these empty cells should be taken into account when we will further aggregate the cube for the production of maps or timelines.

As an example, we present below an extraction of the hypercube of titles (order) of news during the month of september 2015 (when = "2015-09-01") by the french newspaper La Croix (who) about Syria and Hungary (where1 and where2) for the topic of human mobility (what). We have normally 1 x 1 x 1 x 2 x 2 x 4 = 16 possibilities of cells as the dimensions of order, who and time are fixed. But only 10 possibilities out of 16 are realized.

```{r hypercube_sample, echo = FALSE}
sel<-hyper$month


sel<-sel[  order %in% c(1) &
           who=="fr_FRA_lacroi" & 
           when==as.Date("2015-09-01") & 
           where1 %in% c("SYR","HUN") & 
           where2 %in% c("SYR","HUN") 
           ,]
kable(sel[order(-news),])

```

In the most of cases, the news just mention one single country.
In the case of Syria, 51.4 news does not mention the topic, 11.3 mentions "refugees", 3.25 "migrants". In the case of Hungary, 3.5 news does not mention the topic, 16.3 mentions "migrants", 2.3 "refugees". The two countries are associated only one time about a news about asylum seekers where a third country was mentioned, which explains the weight of 0.111 = 1/9.


# EXPLORATION AND VISUALIZATION FONCTIONS

The five dimensions of analysis could be analysed through different aggregation of the dimensions of the hypercubes, leading to different tables
authorizing different modes of visualization. Each function is named according to the dimensions that it is built for.
For the the `what` dimension the different subtopics are never analyzed simultaneously. The default option is the aggregation of all the subtopics.

## WHY : DEFINITION OF THE RESEARCH QUESTION

The researcher interested in a topic (WHAT) can develop two mains strategies to explore the topic through the computation and the visualization of two different indicators :

- The *salience indicator*, defined as the ratio $p_{obs}/p_{est}$ between observed and predicted number of news related to the topic, is ideal for an inductive approach. It is similar to *specialization indexes* used in econometric model and very useful at the initial stage to identify under- or over-representations of the topic of interest during the time units of observation.

- The *p.value* of a chi-2 test applied to the hypothesis of independence is more adapted to a deductive or hypthetico-deductive approach, to improve hypothetisis like the suspition of under- or over-representations revealed by a first observation using the salience index. 
Using an unilateral test ($H_0 : p_{obs} > p_{est}$) we obtain a p-value which can be interpreted as a normalized index of deviation define on $[0,1]$ (with the tool it is possible to interpret the value greater than 0.99 as significant positive exceptions and the values lower than 0.01 as significant negative exceptions).  

All the functions described below will therefore use a statistical test which identify for which value it is possible to say that the proportion of news related to a topic is significantly lower or greater to a reference value. The R function `testchi2` receive as input a table with three columns:

- **trial** : the number of textual units where the topic can be identified: the number of title or sentences that has been collected. The number of trial can be greater than the number of news when we combine the title and the sentences of the description. 
- **success** : the number of textual units where the topic of interest has been found.: a sum of weights than are not necessarily integers, the values are rounded to zero before computation of the test.
- **null.value** : the expected probability of success (i.e. the expected value of the ratio of success divided by trial). In the one-dimension cases, the null.value is equal to a unique value corresponding to the mean probability of success for all sources during the period of observation. But in the case of multi-dimensional analysis, the null.value can be segmented by one of the dimension of interest. 

Once the three variables are fixed, the function will be applied according to a set of three parameters :

- `mintest`: the minimum value of the estimated number of success needed for the chi-square test in optimal statistical conditions. The literature suggest a threshold value of 5 but the user can decide to reduce or increase this threshold. The test will be made if and only if the threshold is reached. We do not recommend to use low threshold: it will generate error messages by using the `prop.test()` function and increase the time need for the computation procedure. 
- `cut-breaks` and `cut_names`are linked parameters used to define qualitative interpretations of the p-value. In our example, we are interested in both value located below or above the threshold. For this reason we will use a double scale of significance  in both directions around the expected probability. 

The result of the procedure is the addition of four columns to the initial table :

- **estimate** is the observed proportion of the news related to the topic (i.e. ratio of success divided by trial) which has to be compared to the reference value,
- **salience** is the ratio between estimate and null.value,
- **chi2** is the value of the chi-square test of the relation between expected and observed success,
- **p.value** is the result of the unilateral test $p_{observed} < p_{estimated}$ with 1 degree of freedom. 

- NB.1: When the threshold `minsamp` is not reached, the columns **estimate** and **salience** are filled with missing values. The aim is to avoid interpretation of variation of proportion that are not based on a reasonable sample of news. The default value for this parameter is 20| .

- NB.2 : When the threshold `minsamp` and `mintest` are not reached together, the columns **chi2** and **p.value** are filled with missing values. The aim is to avoid realization of test that do not fulfill the statistical conditions according to the rules of the art.  The dafault value for this parameter is 5. 


### Function

```{r, echo = TRUE}
#### ---------------- testchi2 ----------------
#' @title  Compute the average salience of the topic and test significance of deviation
#' @name what
#' @description create a table and graphic of the topic
#' @param tabtest a table with variable trial, success and null.value
#' @param minsamp : Threshold of sample size requested for salience computation
#' @param mintest : Threshold of estimated value requested for chi-square test


testchi2<-function(tabtest=tabtest,
                   minsamp = 20,
                   mintest = 5) 
  {
  tab<-tabtest
  n<-dim(tab)[1]
  
  # Compute salience if sample size sufficient (default : N>20)
  tab$estimate <-NA
  tab$salience <-NA
  tab$chi2<-NA
  tab$p.value<-NA
  if (tab$trial > minsamp){ tab$estimate<-round(tab$success/tab$trial,5)
                      tab$salience<-tab$estimate/tab$null.value
  
         # Chi-square test if estimated value sufficient (default : Nij* > 5)

           for (i in 1:n) {
               if(tab$trial[i]*tab$null.value[i]>=mintest) {  
                    test<-prop.test(x=tab$success[i],n=tab$trial[i], p=tab$null.value[i], 
                                alternative = "less")
                    tab$chi2[i]<-round(test$statistic,2)
                    tab$p.value[i]<-round(test$p.value,5)
                 } 
                }
  }
  return(tab)
}
```


```{r}
tab<-hyper$week
tab$what <-tab$what !="_no_"
tab<-tab[is.na(when)==F,]
tab<-tab[,list(trial = sum(news),success=round(sum(news*what),0)),by = list(who)]
ref <-round(sum(tab$success)/sum(tab$trial),4)
tab$null.value<-ref


res<-testchi2(tabtest=tab,
                   minsamp = 100,
                   mintest = 5) 
kable(res)

```






## WHO.WHAT


### Function

```{r who.what function, echo = TRUE}

#### ---------------- who.what ----------------
#' @title  visualize variation of the topic between media
#' @name who.what
#' @description create a table of variation of the topic by media
#' @param hc an hypercube prepared as data.table
#' @param test : visualize test (TRUE) or salience (FALSE)
#' @param minsamp : Threshold of sample size requested for salience computation
#' @param mintest sample size of estimate for chi-square test (default = 5)
#' @param title Title of the graphic


who.what <- function (hc = hypercube,
                      test = FALSE,
                      minsamp = 20,
                      mintest = 5,
                      title = "Who says What ?")
{
  
  tab<-hc
{tab$what <-tab$what !="_no_"}
  
  tab<-tab[,list(trial = sum(news),success=round(sum(news*what),0)),by = list(who)]
  ref <-round(sum(tab$success)/sum(tab$trial),4)
  tab$null.value<-ref
  
  tab<-testchi2(tabtest=tab,
                minsamp = minsamp,
                mintest = mintest)
  

  
    if (test==FALSE) {tab$index =tab$salience
                     tab<-tab[tab$trial > minsamp,]
                     mycol<-brewer.pal(7,"YlOrRd")
                     } 
               else {tab$index=tab$p.value
                     tab<-tab[tab$trial*tab$null.value>mintest,]
                      mycol<-rev(brewer.pal(7,"RdYlBu"))
                      mycol[4]<-"lightyellow"
                     }
  
  p <- plot_ly(tab,
               x = ~who,
               y = ~estimate*100,
               color= ~index,
               colors= mycol,
               hoverinfo = "text",
               text = ~paste('Source: ',who,
                             '<br /> Total news  : ', round(trial,0),
                             '<br /> Topic news : ', round(success,0),
                             '<br /> % observed  : ', round(estimate*100,2),'%',
                             '<br /> % estimated : ', round(null.value*100,2),'%',
                             '<br /> Salience : ', round(salience,2),  
                             '<br /> p.value : ', round(p.value,4)),
               type = "bar")  %>%
    layout(title = title,
           yaxis = list(title = "% news"),
           barmode = 'stack')
  
  output<-list("table" = tab, "plotly" =p)
  
  return(output)
  
}


```


### Example

```{r who.what example}
hc<-hyper$year

res_who_what<- who.what(hc=hc, 
                        test = FALSE,
                        minsamp = 20,
                        mintest = 5,
                        title = "Migrant news by media - Salience")
res_who_what$plotly %>% layout(autosize = F, width = 800, height = 500)

res_who_what<- who.what(hc=hc, 
                        test = TRUE,
                        minsamp = 20,
                        mintest = 5,
                        title = "Migrant news by media - Significance")
res_who_what$plotly %>% layout(autosize = F, width = 800, height = 500)

```

## WHEN.WHAT


### Function

```{r when.what function, echo = TRUE}


#### ---------------- when.what ----------------
#' @title  visualize variation of the topic through time
#' @name when.what
#' @description create a table of variation of the topic by media
#' @param test : visualize test (TRUE) or salience (FALSE)
#' @param minsamp : Threshold of sample size requested for salience computation
#' @param mintest sample size of estimate for chi-square test (default = 5)
#' @param title Title of the graphic


when.what <- function (hc = hypercube,
                      test = FALSE,
                      minsamp = 20,
                      mintest = 5,
                      title = "Who says What ?")
{
  
  tab<-hc
{tab$what <-tab$what !="_no_"}
  
  tab<-tab[,list(trial = sum(news),success=round(sum(news*what),0)),by = list(when)]
  ref <-round(sum(tab$success)/sum(tab$trial),4)
  tab$null.value<-ref
  
  tab<-testchi2(tabtest=tab,
                minsamp = minsamp,
                mintest = mintest)
  
    if (test==FALSE) {tab$index =tab$salience
                     tab<-tab[tab$trial > minsamp,]
                     mycol<-brewer.pal(7,"YlOrRd")
                     } 
               else {tab$index=tab$p.value
                     tab<-tab[tab$trial*tab$null.value>mintest,]
                      mycol<-rev(brewer.pal(7,"RdYlBu"))
                      mycol[4]<-"lightyellow"
                     }
  
  
  p <- plot_ly(tab,
               x = ~as.character(when),
               y = ~estimate*100,
               color= ~index,
               colors= mycol,
               hoverinfo = "text",
               text = ~paste('Time: ',when,
                             '<br /> Total news  : ', round(trial,0),
                             '<br /> Topic news : ', round(success,0),
                             '<br /> % observed  : ', round(estimate*100,2),'%',
                             '<br /> % estimated : ', round(null.value*100,2),'%',
                             '<br /> Salience : ', round(salience,2),  
                             '<br /> p.value : ', round(p.value,4)),
               type = "bar")  %>%
    layout(title = title,
           yaxis = list(title = "% news"),
           barmode = 'stack')
  
  output<-list("table" = tab, "plotly" =p)
  
  return(output)
  
}
```

### Example

```{r when.what example}
hc<-hyper$week
hc[is.na(hc$when)==F,]
res_when_what<- when.what(hc=hc, 
                          test=FALSE,
                          minsamp=20,
                          mintest=5,
                          title = "Migrant news by month - Salience")
res_when_what$plotly %>% layout(autosize = F, width = 800, height = 500)

res_when_what<- when.what(hc=hc, 
                          test=TRUE,
                          minsamp=20,
                          mintest=5,
                          title = "Migrant news by month - Significance")
res_when_what$plotly %>% layout(autosize = F, width = 800, height = 500)
```


## WHERE.WHAT


### Function

```{r where.what function, echo = TRUE}

#### ---------------- where.what ----------------
#' @title  visualize spatialization of the topic 
#' @name where.what
#' @description create a table of variation of the topic by media
#' @param hc an hypercube prepared as data.table
#' @param test : visualize test (TRUE) or salience (FALSE)
#' @param minsamp : Threshold of sample size requested for salience computation
#' @param mintest sample size of estimate for chi-square test (default = 5)
#' @param map a map with coordinates in lat-long
#' @param proj a projection accepted by plotly
#' @param title Title of the graphic


where.what <- function (hc = hypercube,
                        test = FALSE,
                       minsamp = 20,
                       mintest = 5,
                       map = world_ctr,
                       proj = 'azimuthal equal area',
                       title = "Where said What ?")
{

tab<-hc
tab$what <-tab$what !="_no_"

tab<-tab[,list(trial = round(sum(news),0),success=round(sum(news*what),0)),by = list(where1)]
ref <-round(sum(tab$success)/sum(tab$trial),4)
tab$null.value<-ref

  tab<-testchi2(tabtest=tab,
                minsamp = minsamp,
                mintest = mintest)



tab<-tab[order(-chi2),]



    if (test==FALSE) {tab$index =tab$salience
                     tab<-tab[tab$trial > minsamp,]
                     mycol<-brewer.pal(7,"YlOrRd")
                     } 
               else {tab$index=tab$p.value
                     tab<-tab[tab$trial*tab$null.value>mintest,]
                      mycol<-rev(brewer.pal(7,"RdYlBu"))
                      mycol[4]<-"lightyellow"
                     }
  

map<-merge(map,tab,all.x=T,all.y=F,by.x="ISO3",by.y="where1")
  


#map2<-map[is.na(map$pct)==F,]
#map2<-st_centroid(map2)
#map2<-st_drop_geometry(map2)


g <- list(showframe = TRUE,
          framecolor= toRGB("gray20"),
          coastlinecolor = toRGB("gray20"),
          showland = TRUE,
          landcolor = toRGB("gray50"),
          showcountries = TRUE,
          countrycolor = toRGB("white"),
          countrywidth = 0.2,
         projection = list(type = proj))



p<- plot_geo(map)%>%
  add_markers(x = ~lon,
              y = ~lat,
              sizes = c(0, 250),
              size = ~success,
#             color= ~signif,
              color = ~index,
              colors= mycol,
              hoverinfo = "text",
              text = ~paste('Location: ',NAME,
                             '<br /> Total news  : ', round(trial,0),
                             '<br /> Topic news : ', round(success,0),
                             '<br /> % observed  : ', round(estimate*100,2),'%',
                             '<br /> % estimated : ', round(null.value*100,2),'%',
                             '<br /> Salience : ', round(salience,2),  
                             '<br /> p.value : ', round(p.value,4))) %>%
  
  layout(geo = g,
         title = title)

p


output<-list("table" = tab, "plotly" =p)

return(output)

}

```


### Example

```{r where.what example}
map<-readRDS("map/world_ctr_4326.Rdata")
hc<-hyper$year
res_where_what<- where.what(hc=hc,
                            test=FALSE,
                            map = map, 
                            mintest =5,
                            title = "Migrant news by location - Salience")
res_where_what$plotly %>% layout(autosize = F, width = 800, height = 500)
res_where_what<- where.what(hc=hc,
                            test=TRUE,
                            map = map, 
                            mintest =5,
                            title = "Migrant news by location - Significance")
res_where_what$plotly %>% layout(autosize = F, width = 800, height = 500)
```


## WHEN.WHO.WHAT


### Function

```{r when.who.what function, echo=TRUE}

#### ---------------- when.who.what ----------------
#' @title  visualize variation of the topic by media through time
#' @name when.who.what
#' @description create a table of variation of the topic by media through time
#' @param hc an hypercube prepared as data.table
#' @param test : visualize test (TRUE) or salience (FALSE)
#' @param minsamp : Threshold of sample size requested for salience computation
#' @param mintest sample size of estimate for chi-square test (default = 5)
#' @param title Title of the graphic


when.who.what <- function (hc = hypercube,
                           test = FALSE,
                           minsamp = 20,
                           mintest = 5,
                           title = "What by Whom and When ?")
{

tab<-hc
tab$what <-tab$what !="_no_"
tab<-tab[is.na(when)==F,]



tab<-tab[,list(trial = sum(news),success=round(sum(news*what),0)),by = list(when,who)]
ref<-tab[,list(null.value = round(sum(success)/sum(trial),4)), by = list(who)]
tab<-merge(tab,ref,by="who")

  tab<-testchi2(tabtest=tab,
                minsamp = minsamp,
                mintest = mintest)


    if (test==FALSE) {tab$index =tab$salience
                     tab<-tab[tab$trial > minsamp,]
                     mycol<-brewer.pal(7,"YlOrRd")
                     } 
               else {tab$index=tab$p.value
                     tab<-tab[tab$trial*tab$null.value>mintest,]
                      mycol<-rev(brewer.pal(7,"RdYlBu"))
                      mycol[4]<-"lightyellow"
                     }



 p <- plot_ly(tab,
  x = ~when,
  y = ~who,
  z= ~index,
  sizes = c(0, 250),
  size = ~success,
  colors= mycol,
  hoverinfo = "text",
  text = ~paste('Date: ',when,
                '<br> Media: ',who,
                '<br /> Total news  : ', round(trial,0),
                '<br /> Topic news : ', round(success,0),
                '<br /> % observed  : ', round(estimate*100,2),'%',
                '<br /> % estimated : ', round(null.value*100,2),'%',
                '<br /> Salience : ', round(salience,2),  
                '<br /> p.value : ', round(p.value,4)),
#  name = ~tags,
  type = "heatmap")  %>%
  layout(title = title,
         yaxis = list(title = "media"),
         xaxis = list(title = "time"))
p



output<-list("table" = tab, "plotly" =p)

return(output)

}
```


### Example

```{r when.who.what example}
hc<-hyper$month
res_when_who_what<- when.who.what(hc=hc,
                                  test = FALSE,
                                  minsamp = 20,
                                  mintest = 5,
                                  title = "Migrant news through time by media - Salience")
res_when_who_what$plotly %>% layout(autosize = F, width = 800, height = 500)

res_when_who_what<- when.who.what(hc=hc,
                                  test = TRUE,
                                  minsamp = 20,
                                  mintest = 5,
                                  title = "Migrant news through time by media - Significance")
res_when_who_what$plotly %>% layout(autosize = F, width = 800, height = 500)
```


## WHERE.WHO.WHAT


### Function

```{r where.who.what function, echo=TRUE}

#### ---------------- where.who.what ----------------
#' @title  visualize variation of the topic by location and media
#' @name where.who.what
#' @description create a table of variation of the topic by location and media
#' @param hc an hypercube prepared as data.table
#' @param maxloc maximum number of location
#' @param test : visualize test (TRUE) or salience (FALSE)
#' @param minsamp : Threshold of sample size requested for salience computation
#' @param mintest sample size of estimate for chi-square test
#' @param title Title of the graphic


where.who.what <- function (hc = hypercube,
                           maxloc = 15,
                           test = FALSE,
                           minsamp=20,
                           mintest = 5,
                           title = "What by Whom and Where ?")
{

tab<-hc
tab$what <-tab$what !="_no_"

tab<-tab[,list(trial = sum(news),success=round(sum(news*what),0)),by = list(where1,who)]
ref<-tab[,list(null.value = round(sum(success)/sum(trial),4)), by = list(who)]
tab<-merge(tab,ref,by="who")

# selection
sel<-tab[,list(success = sum(success)), by = list(where1)]
sel<-sel[order(-success),]
sel<- sel[1:maxloc,]
tab<-tab[where1 %in% sel$where1,]
tab$trial<-round(tab$trial,0)
tab$success<-round(tab$success,0)


  tab<-testchi2(tabtest=tab,
                minsamp = minsamp,
                mintest = mintest)


    if (test==FALSE) {tab$index =tab$salience
                     tab<-tab[tab$trial > minsamp,]
                     mycol<-brewer.pal(7,"YlOrRd")
                     } 
               else {tab$index=tab$p.value
                     tab<-tab[tab$trial*tab$null.value>mintest,]
                      mycol<-rev(brewer.pal(7,"RdYlBu"))
                      mycol[4]<-"lightyellow"
                     }




 p <- plot_ly(tab,
  x = ~where1,
  y = ~who,
  z= ~index,
  sizes = c(0, 250),
  size = ~success,
  colors= mycol,
  hoverinfo = "text",
  text = ~paste('Location: ',where1,
                '<br>Media: ',who,
                '<br /> Total news  : ', round(trial,0),
                '<br /> Topic news : ', round(success,0),
                '<br /> % observed  : ', round(estimate*100,2),'%',
                '<br /> % estimated : ', round(null.value*100,2),'%',
                '<br /> Salience : ', round(salience,2),  
                '<br /> p.value : ', round(p.value,4)),
#  name = ~tags,
  type = "heatmap")  %>%
  layout(title = title,
         yaxis = list(title = "Host media"),
         xaxis = list(title = "Guest countries"))




output<-list("table" = tab, "plotly" =p)

return(output)

}
```


### Example

```{r where.who.what example}
hc<-hyper$year
res_where_who_what<- where.who.what(hc= hc,
                                    maxloc= 20,
                                    test=FALSE,
                                    minsamp=20,
                                    mintest=5,
                                    title = "Migrant news by media and location - Salience")
res_where_who_what$plotly %>% layout(autosize = F, width = 800, height = 500)

res_where_who_what<- where.who.what(hc= hc,
                                    maxloc= 20,
                                    test=TRUE,
                                    minsamp=20,
                                    mintest=5,
                                    title = "Migrant news by media and location- Significance")
res_where_who_what$plotly %>% layout(autosize = F, width = 800, height = 500)
```


## WHEN.WHERE.WHAT


### Function

```{r when.where.what function, echo=TRUE}
#### ---------------- when.where.what ----------------
#' @title  visualize variation of the topic by location through time
#' @name when.where.what
#' @description create a table of variation of the topic by location through time
#' @param hc an hypercube prepared as data.table
#' @param maxloc maximum number of location
#' @param test : visualize test (TRUE) or salience (FALSE)
#' @param minsamp : Threshold of sample size requested for salience computation
#' @param mintest sample size of estimate for chi-square test
#' @param title Title of the graphic


when.where.what <- function (hc = hypercube,
                           maxloc = 15,
                           test = FALSE,
                           minsamp=20,
                           mintest = 5,
                           title = "What, Where and When ?")
{

tab<-hc
tab$what <-tab$what !="_no_"

tab<-tab[,list(trial = sum(news),success=round(sum(news*what),0)),by = list(where1,when)]
ref<-tab[,list(null.value = round(sum(success)/sum(trial),4)), by = list(where1)]
tab<-merge(tab,ref,by="where1")

# selection
sel<-tab[,list(success = sum(success)), by = list(where1)]
sel<-sel[order(-success),]
sel<- sel[1:maxloc,]
tab<-tab[where1 %in% sel$where1,]
tab$trial<-round(tab$trial,0)
tab$success<-round(tab$success,0)

  tab<-testchi2(tabtest=tab,
                minsamp = minsamp,
                mintest = mintest)


    if (test==FALSE) {tab$index =tab$salience
                     tab<-tab[tab$trial > minsamp,]
                     mycol<-brewer.pal(7,"YlOrRd")
                     } 
               else {tab$index=tab$p.value
                     tab<-tab[tab$trial*tab$null.value>mintest,]
                      mycol<-rev(brewer.pal(7,"RdYlBu"))
                      mycol[4]<-"lightyellow"
                     }



 p <- plot_ly(tab,
  x = ~when,
  y = ~where1,
  z= ~index,
  sizes = c(0, 250),
  size = ~success,
  colors= mycol,
  hoverinfo = "text",
  text = ~paste('Location: ',where1,
                '<br>Date: ',when,
                '<br /> Total news  : ', round(trial,0),
                '<br /> Topic news : ', round(success,0),
                '<br /> % observed  : ', round(estimate*100,2),'%',
                '<br /> % estimated : ', round(null.value*100,2),'%',
                '<br /> Salience : ', round(salience,2),  
                '<br /> p.value : ', round(p.value,4)),
#  name = ~tags,
  type = "heatmap")  %>%
  layout(title = title,
         yaxis = list(title = "Guest countries"),
         xaxis = list(title = "Time"))

p

output<-list("table" = tab, "plotly" =p)

return(output)
}
```


### Example

```{r when.where.what example}
hc<-hyper$month
res_when_where_what<- when.where.what(hc=hc,
                                    maxloc= 15,
                                    test=FALSE,
                                    minsamp=20,
                                    mintest=5,
                                    title = "Migrant news by location through time - Salience")
res_when_where_what$plotly %>% layout(autosize = F, width = 800, height = 500)
res_when_where_what<- when.where.what(hc=hc,
                                    maxloc= 15,
                                    test=TRUE,
                                    minsamp=20,
                                    mintest=5,
                                    title = "Migrant news by location through time - Significance")
res_when_where_what$plotly %>% layout(autosize = F, width = 800, height = 500)

```




## WHERE.WHERE.WHAT


### Function

```{r where.where.what function, echo=TRUE}

#### ---------------- where.where.what ----------------
#' @title  visualize variation of the topic by co-location 
#' @name where.where.what
#' @description create a table of variation of the topic by co-location
#' @param hc an hypercube prepared as data.table
#' @param test : visualize test (TRUE) or salience (FALSE)
#' @param minsamp : Threshold of sample size requested for salience computation
#' @param mintest minimum expected size of news for test
#' @param minedge minimum news with topic by edge
#' @param minnode minimum news with topic by node
#' @param title Title of the graphic


where.where.what <- function (hc = hypercube,
                           test=FALSE,
                           minsamp = 20,
                           mintest = 5,
                           minedge = 2,
                           minnode = 10,
                           title = "What, Where and Where ?")
{

  #test...

  
  
  
tab<-hc
tab$what <-tab$what !="_no_"

# Palette


# Create edges
tab1<-tab[where1 !=where2,list(trial = sum(news),success=round(sum(news*what),0)),by = list(where1,where2)]
tab1$null.value<-sum(tab1$success)/sum(tab1$trial)

tab1<-testchi2(tabtest=tab1,
              minsamp = minsamp,
              mintest = mintest)

    if (test==FALSE) {tab1$index =tab1$salience
                     tab1<-tab1[tab1$trial > minsamp,]
                     mycol<-brewer.pal(7,"YlOrRd") }  else
                       {tab1$index=tab1$p.value
                     tab1<-tab1[tab1$trial*tab1$null.value>mintest,]
                      mycol<-rev(brewer.pal(7,"RdYlBu"))
                      mycol[4]<-"lightyellow"
                     }




tab1<-tab1[order(success),]

# Create nodes
tab2<-tab[where1 !=where2,list(trial = sum(news),success=round(sum(news*what),0)),by = list(where1)]
tab2$null.value<-sum(tab2$success)/sum(tab2$trial)
tab2<-testchi2(tabtest=tab2,
                minsamp = minsamp,
                mintest = mintest)

    if (test==FALSE) {tab2$index =tab2$salience
                     tab2<-tab2[tab2$trial > minsamp,]
                     mycol<-brewer.pal(7,"YlOrRd") }  else
                       {tab2$index=tab2$p.value
                     tab2<-tab2[tab2$trial*tab2$null.value>mintest,]
                      mycol<-rev(brewer.pal(7,"RdYlBu"))
                      mycol[4]<-"lightyellow"
                     }

tab2<-tab2[order(success),]




# Build tibble graph object
tib_g=tidygraph::tbl_graph(nodes=tab2,edges=tab1) 

# filter
sel_tib_g <-tib_g %>% activate(edges) %>%
                      filter(success > minedge) %>%
                      activate(nodes) %>%
                      filter(success > minnode) %>%
                      mutate(isol = node_is_isolated()) %>%
                      filter(isol == FALSE)

## Create a ggraph layout
g=sel_tib_g %>% 
   ggraph(layout="stress")

# visualize
gg<-g + geom_edge_link(aes(edge_width=success, edge_colour = index),
                   alpha = 0.3 , show.legend=c(TRUE,FALSE,FALSE, FALSE,FALSE)) +
        scale_edge_colour_gradientn(colors = mycol)+
   geom_node_point(aes(colour = index, size=success), 
                  alpha=0.6) +
        scale_color_gradientn(colors =mycol)+
  geom_node_label(aes(label = where1, size = 2*sqrt(success)),alpha =1,label.size=0.1,show.legend = FALSE)


output<-list("table" = tab1, "plot" =gg)

return(output)
}
```

### Example
```{r where.where.what example}

res_where_where_what<- where.where.what(hc=hyper$year,
                                        test=FALSE,
                                         mintest = 5,
                                         minsamp = 10,
                                         minedge = 1,
                                         minnode = 2,
                                         title = "Network of news about topic - Salience")
res_where_where_what$plot 

res_where_where_what<- where.where.what(hc=hyper$year,
                                        test=TRUE,
                                         mintest = 5,
                                         minsamp = 10,
                                         minedge = 1,
                                         minnode =2,
                                         title = "Network of news about topic - Significance")
res_where_where_what$plot 


```