---
title: "MODELS USED IN FRONTIER'S PAPER (V2)"
author: "Claude Grasland, Université de Paris"
subtitle: Version 1.0 - August 2021
output:
  pdf_document: default
  html_document: default
---



```{r setup, echo = FALSE, message = FALSE, warning = FALSE, error = FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, error = FALSE)
library(sf)
library(plotly)
library(RColorBrewer)
library(data.table)
library(DT)
library(dplyr)
library(cartography)
library(knitr)
library(tidygraph)
library(ggraph)
library(ggplot2)
library(stargazer)
#library(reldist)
#library(quanteda)

#load hypercube by weeks
hc <- readRDS("data/hc_EUR_foreign_mobil_week.Rdata")
hc<-hc[hc$when>as.Date("2013-12-31"),]
hc<-hc[hc$when<as.Date("2020-01-01"),]
hc<-hc[is.na(hc$when)==FALSE,]
hc<-hc[!hc$who %in% c("es_ESP_vangua","en_GBR_mirror"),]
hc$when<-substr(as.character(hc$when),1,7)
hcmigr<-hc

hc <- readRDS("data/hc_EUR_foreign_border_week.Rdata")
hc<-hc[hc$when>as.Date("2013-12-31"),]
hc<-hc[hc$when<as.Date("2020-01-01"),]
hc<-hc[is.na(hc$when)==FALSE,]
hc<-hc[!hc$who %in% c("es_ESP_vangua","en_GBR_mirror"),]
hc$when<-substr(as.character(hc$when),1,7)
hcbord<-hc

hc <- readRDS("data/hc_EUR_foreign_pandemic_week.Rdata")
hc<-hc[hc$when>as.Date("2013-12-31"),]
hc<-hc[hc$when<as.Date("2020-01-01"),]
hc<-hc[is.na(hc$when)==FALSE,]
hc<-hc[!hc$who %in% c("es_ESP_vangua","en_GBR_mirror"),]
hc$when<-substr(as.character(hc$when),1,7)
hcpand<-hc


# Load functions

source("pgm/hypernews_functions_V4.R")
```

In the versio 2, we eliminate two newspapers that are problematic (Daily Mirror & La Vanguardia) and we realize the test of hypothesis two at month level rather than week. 


## Model I : National effect / What =f(who-what)

Our first model will test the existence of comparable levels of interest for geopolitical topics in the different media of different countries. It is not really a research hypothesis but rather a control on the joint choice of media and topics linked to the choice of keywords. We assume therefore H1 : 

- **H1 : The global level of interest of media for a geopolitical topic is not related to the country where the media is located.**







```{r}
hc<-hcbord
# table
dt<-who.what(hc)$table
dt$country<-substr(dt$who,4,6)
dt1<-data.frame(media=dt$who,host.country = dt$country,salience=dt$estimate*100, topic="Borders", sample.size=dt$trial)


hc<-hcmigr
# table
dt<-who.what(hc)$table
dt$country<-substr(dt$who,4,6)
dt2<-data.frame(media=dt$who,host.country = dt$country,salience=dt$estimate*100, topic="Migrants & refugees", sample.size=dt$trial)



hc<-hcpand
# table
dt<-who.what(hc)$table
dt$country<-substr(dt$who,4,6)
dt3<-data.frame(media=dt$who,host.country = dt$country,salience=dt$estimate*100, topic="Pandemics", sample.size=dt$trial)


dt<-rbind(dt1,dt2,dt3)

# Plot
p<-  ggplot(dt,aes(x=host.country,y=salience, size=sample.size))+geom_point()+facet_wrap(~topic) + scale_x_discrete("Host countries")+
    scale_y_log10("% of foreign news", breaks = c(0.25, 0.5,1, 2, 4))
p

ggsave(filename="hyp11bis.pdf",plot=p, width = 8, height =4)
?geom_point()

# Model
dt1$border<-dt1$salience
mod1 <-lm(formula = border~host.country,
           data=dt1)
summary(mod1)

dt2$migrants.refugees<-dt2$salience
mod2 <-lm(formula = migrants.refugees~host.country,
           data=dt2)
summary(mod2)

dt3$pandemics<-dt3$salience
mod3 <-lm(formula = pandemics~host.country,
           data=dt3)
summary(mod3)

stargazer(mod1,mod2,mod3)

summary(dt3$salience)
sd(dt3$salience)
sd(dt3$salience)/mean(dt3$salience)


```




## Model 2 : National effect / What = (When, Who-Where)

Our second hypothesis is related to the temporal dimension of the geopolitical agenda. We assume that, all things being equal with the global level of interest for a geopolitical topic, the distribution of the period of interest is influenced by national agenda which produce significant differences in the timeline of the topic over the period of observation. H2 can therefore be formulated as :

- **H2 : The variation of the salience of a geopolitical topic through time is influenced by national agenda. The correlation of the timeline is higher between media of the same country than b etween media of different country**

To verify H2, we build a correlation matrix between the levels of salience of the 20 media outlets (using pairwise complete observation when missing data are present) and we compare the distribution of the level of correlation for media located in the same country and media located in different countries with variance analysis and Fisher test.

### Graphic

```{r Model H1.2 - graphic}

hc<-hcbord
dt<-when.who.what(hc)$table
dt$country<-substr(dt$who,4,6)
dt$week<-as.character(dt$when)
x<-data.frame(dt[,c("who","country","estimate", "when")])
y<-reshape2::dcast(x,formula = when~who,value.var = "estimate")
m<-y[,-1]
row.names(m)<-y$when
cor<-cor(m,method ="pearson",use="pairwise.complete.obs")
col<-melt(cor)
col<-col[col$Var1!=col$Var2,]
col$host.country<-as.factor(substr(col$Var1,4,6)==substr(col$Var2,4,6))
levels(col$host.country)<-c("Different","Same")
col1<-col
col1$topic<-"Borders"


hc<-hcmigr
dt<-when.who.what(hc)$table
dt$country<-substr(dt$who,4,6)
dt$week<-as.character(dt$when)
x<-data.frame(dt[,c("who","country","estimate", "when")])
y<-reshape2::dcast(x,formula = when~who,value.var = "estimate")
m<-y[,-1]
row.names(m)<-y$when
cor<-cor(m,method ="pearson",use="pairwise.complete.obs")
col<-melt(cor)
col<-col[col$Var1!=col$Var2,]
col$host.country<-as.factor(substr(col$Var1,4,6)==substr(col$Var2,4,6))
levels(col$host.country)<-c("Different","Same")
col2<-col
col2$topic<-"Migrants & Refugees"


hc<-hcpand
dt<-when.who.what(hc)$table
dt$country<-substr(dt$who,4,6)
dt$week<-as.character(dt$when)
x<-data.frame(dt[,c("who","country","estimate", "when")])
y<-reshape2::dcast(x,formula = when~who,value.var = "estimate")
m<-y[,-1]
row.names(m)<-y$when
cor<-cor(m,method ="pearson",use="pairwise.complete.obs")
col<-melt(cor)
col<-col[col$Var1!=col$Var2,]
col$host.country<-as.factor(substr(col$Var1,4,6)==substr(col$Var2,4,6))
levels(col$host.country)<-c("Different","Same")
col3<-col
col3$topic<-"Pandemics"

pearson<-rbind(col1,col2,col3)
pearson$correlation<-"Bravais-Pearson"

#---------------------------------------

hc<-hcbord
dt<-when.who.what(hc)$table
dt$country<-substr(dt$who,4,6)
dt$week<-as.character(dt$when)
x<-data.frame(dt[,c("who","country","estimate", "when")])
y<-reshape2::dcast(x,formula = when~who,value.var = "estimate")
m<-y[,-1]
row.names(m)<-y$when
cor<-cor(m,method ="spearman",use="pairwise.complete.obs")
col<-melt(cor)
col<-col[col$Var1!=col$Var2,]
col$host.country<-as.factor(substr(col$Var1,4,6)==substr(col$Var2,4,6))
levels(col$host.country)<-c("Different","Same")
col1<-col
col1$topic<-"Borders"


hc<-hcmigr
dt<-when.who.what(hc)$table
dt$country<-substr(dt$who,4,6)
dt$week<-as.character(dt$when)
x<-data.frame(dt[,c("who","country","estimate", "when")])
y<-reshape2::dcast(x,formula = when~who,value.var = "estimate")
m<-y[,-1]
row.names(m)<-y$when
cor<-cor(m,method ="spearman",use="pairwise.complete.obs")
col<-melt(cor)
col<-col[col$Var1!=col$Var2,]
col$host.country<-as.factor(substr(col$Var1,4,6)==substr(col$Var2,4,6))
levels(col$host.country)<-c("Different","Same")
col2<-col
col2$topic<-"Migrants & Refugees"


hc<-hcpand
dt<-when.who.what(hc)$table
dt$country<-substr(dt$who,4,6)
dt$week<-as.character(dt$when)
x<-data.frame(dt[,c("who","country","estimate", "when")])
y<-reshape2::dcast(x,formula = when~who,value.var = "estimate")
m<-y[,-1]
row.names(m)<-y$when
cor<-cor(m,method ="spearman",use="pairwise.complete.obs")
col<-melt(cor)
col<-col[col$Var1!=col$Var2,]
col$host.country<-as.factor(substr(col$Var1,4,6)==substr(col$Var2,4,6))
levels(col$host.country)<-c("Different","Same")
col3<-col
col3$topic<-"Pandemics"

spearman<-rbind(col1,col2,col3)
spearman$correlation<-"Spearman"


dt<-rbind(pearson,spearman)


# Plot
p<-  ggplot(dt,aes(x=host.country,y=value, fill=host.country))+geom_boxplot(show.legend = FALSE)+facet_wrap(~correlation+topic) + scale_x_discrete("Host countries") +
    scale_y_continuous("Correlation of media agenda by week", breaks = c(0,0.2,0.4,0.6,0.8,1))
p

ggsave(filename="hyp12bis.pdf",plot=p, width = 8, height =4)







```

### Model

```{r Model H1.2 - table}
sel<-dt[dt$correlation=="Bravais-Pearson" & dt$topic=="Borders",]
sel$borders.1<-sel$value
mod1a <-lm(formula = borders.1~host.country,
           data=sel)
summary(mod1a)

sel<-dt[dt$correlation=="Spearman" & dt$topic=="Borders",]
sel$borders.2<-sel$value
mod1b <-lm(formula = borders.2~host.country,
           data=sel)
summary(mod1b)


sel<-dt[dt$correlation=="Bravais-Pearson" & dt$topic=="Migrants & Refugees",]
sel$migrants.1<-sel$value
mod2a <-lm(formula = migrants.1~host.country,
           data=sel)
summary(mod2a)

sel<-dt[dt$correlation=="Spearman" & dt$topic=="Migrants & Refugees",]
sel$migrants.2<-sel$value
mod2b <-lm(formula = migrants.2~host.country,
           data=sel)
summary(mod2b)

sel<-dt[dt$correlation=="Bravais-Pearson" & dt$topic=="Pandemics",]
sel$Pandemics.1<-sel$value
mod3a <-lm(formula = Pandemics.1~host.country,
           data=sel)
summary(mod3a)

sel<-dt[dt$correlation=="Spearman" & dt$topic=="Pandemics",]
sel$Pandemics.2<-sel$value
mod3b <-lm(formula = Pandemics.2~host.country,
           data=sel)
summary(mod3b)

stargazer(mod1a, mod2a,mod3a)


stargazer(mod1b, mod2b,mod3b)




```




## Model 3 : National effect / correlation of Geopolitical maps

We turn now to the second dimension of the geopolitical agenda which is related to the choice of foreign countries mentioned in news related to a geopolitcal topic. The fact to use a relatively long period is particularly interesting in this case because it can help us to have a more structural view than an event oriented approach. It is clear for example that all media outlets has mentioned Turkey and Greece in relation with the migrant topic in september 2015. But it is not so obvious that media of each countries has mentioned these countries in relation with migrant *before* or *after* the period of maximal interest. For the same reason, it is not obvious that media of all countries has offered the same coverage of migrant crisis located outside Europe and its neighborhood like the Rohyngias crisis or the wall established by D. Trump. It is the reason why it can be interested to introduce our third hypothesis as follow. 

- **H3 : Over a sufficiently long period of time the choice of foreign countries associated to  geopolitical topic is influenced by national perceptions of newsworthiness. The correlation between the spatial distribution  of coutries mentionned in relation with a topic is higher between media of the same country than between media of different country**


### Graphic

```{r Model H1.3 - graphic}
# define filter for small countries




hc<-hcbord
dt<-hc[what!="_no_",list(tot=sum(news)),list(where1,who)]
y<-reshape2::dcast(dt,formula = where1~who,value.var = "tot",fill = 0)
y<-y[!(y$where1 %in% c("FRA","DEU","GBR","ESP","ITA")),]
m<-y[,-1]
row.names(m)<-y$when
cor<-cor(m,method ="pearson",use="pairwise.complete.obs")
col<-melt(cor)
col<-col[col$Var1!=col$Var2,]
col$host.country<-as.factor(substr(col$Var1,4,6)==substr(col$Var2,4,6))
levels(col$host.country)<-c("Different","Same")
col1<-col
col1$topic<-"Borders"


hc<-hcmigr
dt<-hc[what!="_no_",list(tot=sum(news)),list(where1,who)]
y<-reshape2::dcast(dt,formula = where1~who,value.var = "tot",fill = 0)
y<-y[!(y$where1 %in% c("FRA","DEU","GBR","ESP","ITA")),]
m<-y[,-1]
row.names(m)<-y$when
cor<-cor(m,method ="pearson",use="pairwise.complete.obs")
col<-melt(cor)
col<-col[col$Var1!=col$Var2,]
col$host.country<-as.factor(substr(col$Var1,4,6)==substr(col$Var2,4,6))
levels(col$host.country)<-c("Different","Same")
col2<-col
col2$topic<-"Migrants & Refugees"


hc<-hcpand
dt<-hc[what!="_no_",list(tot=sum(news)),list(where1,who)]
y<-reshape2::dcast(dt,formula = where1~who,value.var = "tot",fill = 0)
y<-y[!(y$where1 %in% c("FRA","DEU","GBR","ESP","ITA")),]
m<-y[,-1]
row.names(m)<-y$when
cor<-cor(m,method ="pearson",use="pairwise.complete.obs")
col<-melt(cor)
col<-col[col$Var1!=col$Var2,]
col$host.country<-as.factor(substr(col$Var1,4,6)==substr(col$Var2,4,6))
levels(col$host.country)<-c("Different","Same")
col3<-col
col3$topic<-"Pandemics"

pearson<-rbind(col1,col2,col3)
pearson$correlation<-"Bravais-Pearson"

#---------------------------------------

hc<-hcbord
dt<-hc[what!="_no_",list(tot=sum(news)),list(where1,who)]
y<-reshape2::dcast(dt,formula = where1~who,value.var = "tot",fill = 0)
y<-y[!(y$where1 %in% c("FRA","DEU","GBR","ESP","ITA")),]
m<-y[,-1]
row.names(m)<-y$when
cor<-cor(m,method ="spearman",use="pairwise.complete.obs")
col<-melt(cor)
col<-col[col$Var1!=col$Var2,]
col$host.country<-as.factor(substr(col$Var1,4,6)==substr(col$Var2,4,6))
levels(col$host.country)<-c("Different","Same")
col1<-col
col1$topic<-"Borders"


hc<-hcmigr
dt<-hc[what!="_no_",list(tot=sum(news)),list(where1,who)]
y<-reshape2::dcast(dt,formula = where1~who,value.var = "tot",fill = 0)
y<-y[!(y$where1 %in% c("FRA","DEU","GBR","ESP","ITA")),]
m<-y[,-1]

mm<-100*m/apply(m,2,sum)
row.names(m)<-y$where1
cor<-cor(m,method ="spearman",use="pairwise.complete.obs")
col<-melt(cor)
col<-col[col$Var1!=col$Var2,]
col$host.country<-as.factor(substr(col$Var1,4,6)==substr(col$Var2,4,6))
levels(col$host.country)<-c("Different","Same")
col2<-col
col2$topic<-"Migrants & Refugees"


hc<-hcpand
dt<-hc[what!="_no_",list(tot=sum(news)),list(where1,who)]
y<-reshape2::dcast(dt,formula = where1~who,value.var = "tot",fill = 0)
y<-y[!(y$where1 %in% c("FRA","DEU","GBR","ESP","ITA")),]
m<-y[,-1]
row.names(m)<-y$when
cor<-cor(m,method ="spearman",use="pairwise.complete.obs")
col<-melt(cor)
col<-col[col$Var1!=col$Var2,]
col$host.country<-as.factor(substr(col$Var1,4,6)==substr(col$Var2,4,6))
levels(col$host.country)<-c("Different","Same")
col3<-col
col3$topic<-"Pandemics"

spearman<-rbind(col1,col2,col3)
spearman$correlation<-"Spearman"


dt<-rbind(pearson,spearman)


# filter small countries



# Plot
p<-  ggplot(dt,aes(x=host.country,y=value, fill=host.country))+geom_boxplot(show.legend = FALSE)+facet_wrap(~correlation+topic) + scale_x_discrete("Host countries") +
    scale_y_continuous("Correlation between geopolitical maps", breaks = c(0,0.2,0.4,0.6,0.8,1), lim=c(0,1))
p

ggsave(filename="hyp13bis.pdf",plot=p, width = 8, height =4)







```

### Model

```{r Model H1.2 - table}
sel<-dt[dt$correlation=="Bravais-Pearson" & dt$topic=="Borders",]
sel$borders.1<-sel$value
mod1a <-lm(formula = borders.1~host.country,
           data=sel)
summary(mod1a)

sel<-dt[dt$correlation=="Spearman" & dt$topic=="Borders",]
sel$borders.2<-sel$value
mod1b <-lm(formula = borders.2~host.country,
           data=sel)
summary(mod1b)


sel<-dt[dt$correlation=="Bravais-Pearson" & dt$topic=="Migrants & Refugees",]
sel$migrants.1<-sel$value
mod2a <-lm(formula = migrants.1~host.country,
           data=sel)
summary(mod2a)

sel<-dt[dt$correlation=="Spearman" & dt$topic=="Migrants & Refugees",]
sel$migrants.2<-sel$value
mod2b <-lm(formula = migrants.2~host.country,
           data=sel)
summary(mod2b)

sel<-dt[dt$correlation=="Bravais-Pearson" & dt$topic=="Pandemics",]
sel$Pandemics.1<-sel$value
mod3a <-lm(formula = Pandemics.1~host.country,
           data=sel)
summary(mod3a)

sel<-dt[dt$correlation=="Spearman" & dt$topic=="Pandemics",]
sel$Pandemics.2<-sel$value
mod3b <-lm(formula = Pandemics.2~host.country,
           data=sel)
summary(mod3b)

stargazer(mod1a, mod2a,mod3a)


stargazer(mod1b, mod2b,mod3b)




```



